---
title: "Soil ECa"
author: "Ujjwal Sigdel"
format:
  html:
    code-fold: false
    embed-resources: true
    toc: true
    number-sections: true
    theme: journal
    #cerulean, journal, quartz, sketchy, slate, morph, superhero, vapor
knitr: 
  opts_chunk: 
    message: false
    warning: false
---

# Objectives  
To wrangle and analysis soil EC data for Corn for Precision Ag class final project

# Setup  
```{r setup }
# Packages
library(dplyr)
library(tidyr)
library(janitor)
library(readr)
library(sf) #vector manipulation
library(ggplot2)
library(viridis)
library(ggthemes)
library(gstat) #IDW
library(stars) #raster manipulation
library(terra) #topographical variables
library(patchwork) #combining multiple plots  
library(tibble) #rownames to column

```

```{r defining map theme}
mytheme_map <- 
  theme_map()+
  theme(legend.position = "right",
        panel.background = element_rect(fill="gray80",
                                        color=NA),
        plot.background = element_rect(fill="gray80",
                                       color=NA),
        legend.title = element_text(size=14),
        legend.text = element_text(size=12))

```

```{r reading elevation data}
eca <- read_csv("../data/Deer Run1 Soil EC Data.csv")

eca
```
# Wrangling  
```{r eca_w}
eca_w <- eca %>% 
  clean_names() %>% 
  dplyr::select(lon = longitude,
                lat = latitude,
                eca30_dsm = shallow_eca,
                eca90_dsm = deep_eca,
                elev_m = elevation_m
                ) %>% 
  st_as_sf(coords = c("lon", "lat")) %>% 
  st_set_crs(value = 4326) %>% 
  st_transform(crs = 6345)


eca_w 
```

# EDA  
```{r eca30 plot}
eca_w %>% 
  ggplot()+
  geom_sf(aes(color = eca30_dsm),
          size = .5)+
  scale_color_viridis_b()+
  mytheme_map

```


```{r eca90 plot}
eca_w %>% 
  ggplot()+
  geom_sf(aes(color = eca90_dsm),
          size = .5)+
  scale_color_viridis_b()+
  mytheme_map

```

# Grid  
Since we are interpolating, we need to create a grid (same as for the yield and terrain exercises) based on the field boundary.    

```{r boundary}
boundary <- read_sf("../data/boundary/DR1 boundary.shp") %>% 
  st_transform(crs = 6345)

boundary
```

```{r grid_r}
# grid in vector format
grid_r <- boundary %>% 
  st_make_grid(cellsize = 10) %>% 
  st_as_sf() %>% 
  st_rasterize(dx = 10,
               dy = 10) %>% 
  st_crop(boundary)

grid_r

ggplot()+
  geom_stars(data = grid_r)+
  mytheme_map

```

# Interpolation
## eca30  
```{r eca30_idw_mod}
eca30_idw_mod <- gstat(formula = eca30_dsm ~1,
                       data = eca_w)

eca30_idw_mod 
```

```{r eca30_idw_pred}
eca30_idw_pred <- predict(eca30_idw_mod, grid_r)  

eca30_idw_pred
```

```{r eca30_map }
eca30_map <- ggplot()+
  geom_stars(data = eca30_idw_pred)+
  scale_fill_viridis_c()+
  geom_sf(data = boundary, fill=NA)+
  labs(fill = "ECa\n0-30 cm\n(ds/m)")+
  mytheme_map

eca30_map
```

## eca90  
```{r eca90_idw_mod}
eca90_idw_mod <- gstat(formula = eca90_dsm ~ 1,
                       data = eca_w)

eca90_idw_mod
```

```{r eca90_idw_pred}
eca90_idw_pred <- predict(eca90_idw_mod, grid_r)

eca90_idw_pred
```

```{r eca90_map }
eca90_map <- ggplot()+
  geom_stars(data = eca90_idw_pred)+
  scale_fill_viridis_c()+
  geom_sf(data=boundary, fill=NA)+
  labs(fill="ECa\n0-90 cm\n(dS/m)")+
  mytheme_map

eca90_map
```

## Plotting all together
```{r plotting all together}
eca30_map + eca90_map

ggsave("../output/soileca.png", 
       width = 6, 
       height = 3)
```

# Grid extraction  
```{r eca30_v}
eca30_v <- eca30_idw_pred %>% 
  st_as_sf(as_points = F, #if F is not done, by default it is true, and it will give the centroids
           merge = F) %>% 
  dplyr::select(eca30_dsm = var1.pred)
eca30_v
```


```{r eca90_v}
eca90_v <- eca90_idw_pred %>%
  st_as_sf(as_points = F, 
           merge = F) %>%
  dplyr::select(eca90_dsm=var1.pred) 

eca90_v
```

# Merging eca layers  
```{r eca_v }
eca_v <- eca30_v %>% 
  st_join(eca90_v,
          join = st_equals,
          left = T)

eca_v
```

# Exporting  
Exporting to file to use later for management zone creation.  
```{r exporting}
write_sf(eca_v,
         "../data/eca_v.geojson",
         delete_dsn = T)

```
