---
title: "Yield monitor wrangling - 2017"
format:
  html:
    code-fold: false
    embed-resources: true
    toc: true
    number-sections: true
    theme: journal
    #cerulean, journal, quartz, sketchy, slate, morph, superhero, vapor
knitr: 
  opts_chunk: 
    message: false
    warning: false
---

# Objectives  

# Setup  
## Loading packages  
```{r setup }
#| message: false

#loading packages (add to the list as we need) 
library(readr)
library(janitor)
library(dplyr)
library(tidyr)
library(ggplot2)
library(sf)
library(viridis)
library(mapview)

```

## Importing data  
```{r yield17 }
yield <- read_csv("../data/Deer Run1 2016 Yield.csv") %>% 
  clean_names() %>% 
  dplyr::select(long, lat, speed_mph, dry_yield) %>% 
  rename(lon = long, 
         lat = lat,
         yield_buac = dry_yield) %>% 
  mutate(yield_kgha = yield_buac*67.25,
         speed_kmh = speed_mph*1.609344) %>% 
  dplyr::select(lon, lat, yield_kgha, speed_kmh)

yield 

```

# EDA  
```{r summary}
summary(yield) 
```


```{r yield distribution - density plot}
ggplot(yield, aes(x=yield_kgha))+
  geom_density()

```

```{r speed distribution - density plot}
ggplot(yield, aes(x=speed_kmh))+
  geom_density()+
  geom_rug()

```

```{r yield map}
ggplot(yield, aes(x= lon,
                    y = lat))+
  geom_point(aes(color = yield_kgha))+
  scale_color_viridis_b()

```

```{r yield restricted map}
yield %>% 
  filter(yield_kgha < 25000) %>% 
ggplot(aes(x= lon,
                    y = lat))+
  geom_point(aes(color = yield_kgha))+
  scale_color_viridis_b()
  

```

```{r yield x speed}
ggplot(yield, aes(x=speed_kmh, 
                    y=yield_kgha))+
  geom_point()+
  geom_smooth()

```


# Wrangling  
```{r yield17_w}
  yield_w <- yield %>% 
  st_as_sf(coords = c("lon", "lat")) %>% 
  st_set_crs(value = 4326) %>% 
  st_transform(crs = 6345) #%>%
  # st_coordinates() %>%
  # as.data.frame() %>%
  # rename(x = X, y = Y) %>%
  # cbind(yield, .) %>% 
  # select(-lat, -lon)
  
  
yield

```

# Maps  
```{r static}
yield_w %>%
  filter(yield_kgha < 25000) %>%
  ggplot()+
  geom_sf(aes(color = yield_kgha), size = .1)+
  scale_color_viridis_b()

```

## Summary  
```{r }
summary(yield_w) 
```

## Raw map
```{r}
yieldmap_raw <- ggplot(yield_w)+
                geom_sf(aes(color = yield_kgha), size = .1)+
                scale_color_viridis_b()

yieldmap_raw
```

## Raw map interactive  
```{r}
#| warning: false

yieldmap_raw_int <- yield_w %>%
  filter(yield_kgha < 25000) %>%
  mapview(zcol = "yield_kgha",
        col.regions = viridis(3),
        alpha.regions = 1,
        cex = .1,
        lwd = 0,
        map.types = 'Esri.WorldImagery'
)

yieldmap_raw_int
```

Today we are applying 5 filters:  

- Min and max yield  
- Min and max speed  
- Position  

# Yield editor - Min/max yield  
## Raw yield distribution  

```{r}
ggplot(data = yield_w, 
       aes(x = yield_kgha))+
  geom_density()+
  geom_rug()
```

## Raw yield summary  
```{r}
summary(yield_w$yield_kgha)
```

## Raw yield quantiles  
```{r}
lowq_yield <- quantile(yield_w$yield_kgha, 0.1)
lowq_yield

upq_yield <- quantile(yield_w$yield_kgha, 0.99) 
upq_yield

iqr <- upq_yield - lowq_yield
iqr

miny <- lowq_yield - (iqr*0.25)
miny

maxy <- upq_yield + (iqr*0.25) 
maxy
```

## Min/max raw yield quantiles plot  
```{r}
ggplot()+
  geom_density(data = yield_w, 
               aes(x = yield_kgha))+
  geom_vline(xintercept = lowq_yield, color = "red", linetype = 2)+
  geom_vline(xintercept = upq_yield, color = "blue", linetype = 2)+
  geom_vline(xintercept = miny, color = "red", linetype = 1)+
  geom_vline(xintercept = maxy, color = "blue", linetype = 1)+
  annotate(geom = 'rect',
           ymin = 0, 
           ymax = Inf,
           xmin = miny,
           xmax = maxy,
           alpha = .4,
           fill = 'green'
             )

```

## Filtering min/max raw yield  
```{r}
yield_w_yf <- yield_w %>% 
  filter(yield_kgha > miny & yield_kgha < maxy)

yield_w_yf
```

Before filtering we had a total of 31,724 data rows.
How many rows now after filtering?

## Min/max yield cleaned summary  
```{r}
summary(yield_w_yf$yield_kgha)
```
All right, all done filtering based on yield min and max.
Now let’s do the same for speed.

# Yield editor - Min/max speed  
## Raw speed distribution  
```{r}
ggplot(data = yield_w, 
       aes(x = speed_kmh))+
  geom_density()+
  geom_rug()
```

## Raw speed summary  
```{r}
summary(yield_w$speed_kmh) 
```

## Raw speed quantiles  
```{r}
lowq_speed <- quantile(yield_w$speed_kmh, 0.08)
lowq_speed

upq_speed <- quantile(yield_w$speed_kmh, 0.99)
upq_speed

iqr <- upq_speed - lowq_speed
iqr

minv <- lowq_speed - (0.18*iqr)
minv

maxv <- upq_speed + (0.18*iqr)
maxv

```

## Min/max raw speed quantiles plot  
```{r}
ggplot()+
  geom_density(data = yield_w, 
               aes(x = speed_kmh))+
  geom_vline(xintercept = lowq_speed, 
             color = "red", 
             linetype = 2)+
  geom_vline(xintercept = upq_speed, 
             color = "blue", 
             linetype = 2)+
  geom_vline(xintercept = minv, 
             color = "red", 
             linetype =1)+
  geom_vline(xintercept = maxv, 
             color = "blue", 
             linetype = 1)+
  annotate(geom = "rect",
           ymin = 0, 
           ymax = Inf,
           xmin = minv, 
           xmax = maxv,
           alpha = .6, 
           fill = "green")

```

## Filtering min/max raw speed  
```{r}
yield17_w_sf <- yield17_w_yf %>% 
  filter(speed_mph > minv & speed_mph < maxv)

yield17_w_sf

```

## Min/max speed cleaned summary  
```{r}
summary(yield17_w_sf$speed_mph)
```

# Yield editor - Position  
Original position filter was based on quantiles too.
Why filter based on boundary?  

## Field boundary  
```{r}
boundary <- read_sf("../../05_digitizing/output/boundary.geojson")

boundary
#mapview(boundary)
```

```{r}
boundary_w <- boundary %>% 
  st_transform(crs = 6345)
  
boundary_w
```

## Field boundary plot  
```{r}
yieldmap_raw +
  geom_sf(data = boundary_w, fill = NA)

```

Let’s create a negative buffer of 20 m.  

## Buffer  
```{r}
buffer <- boundary_w %>% 
  st_buffer(dist = -25) 
  

buffer
```

```{r}
yieldmap_raw + 
  geom_sf(data=boundary_w, 
          fill = NA) + 
  geom_sf(data = buffer, 
          color = "orange", fill = NA, linewidth = 1)
```


## Filtering position  
```{r}
yield17_w_pf <- yield17_w_sf %>% 
  st_intersection(buffer)

yield17_w_pf
```

## Position yield cleaned summary  
```{r}
summary(yield17_w_pf$yield_lbac)
```

# Raw vs Cleaned data  
## Summary statistics  
### Raw yield summary  
```{r}
summary_raw <- yield17_w %>% 
  summarize(n = length(yield_lbac),
            min = min(yield_lbac),
            mean = mean(yield_lbac),
            max = max(yield_lbac),
            sd = sd(yield_lbac),
            cv = (sd/mean)*100
            ) %>% 
  mutate(data = 'raw') %>% 
  dplyr::select(data, everything())

summary_raw
```

### Clean yield summary  
```{r}
summary_filtered <- yield17_w_pf %>%
  summarise(n = length(yield_lbac),
            min = min(yield_lbac),
            mean = mean(yield_lbac),
            max = max(yield_lbac),
            sd = sd(yield_lbac),
            cv = (sd/mean)*100
            ) %>%
  mutate(data = "clean") %>%
  dplyr::select(data, everything())

summary_filtered
```

### Combined yield summary  
```{r}
summary_raw %>%
  bind_rows(summary_filtered)
```

## Density plot  
```{r}
ggplot()+
  geom_density(data = yield17_w, 
               aes(x = yield_lbac), 
               fill = "blue", 
               alpha = .5)+
  geom_density(data = yield17_w_pf,
               aes(x = yield_lbac),
               fill = "green",
               alpha = .4)
```

# Maps  
## Raw yield map  
```{r}
yieldmap_raw

```

## Clean yield map  
```{r}
yieldmap_clean <- ggplot(yield17_w_pf)+
  geom_sf(aes(color = yield_lbac), 
          size = .1)+
  scale_color_viridis_b()

yieldmap_clean
```

## Clean yield map interactive  
```{r}
yieldmap_clean_int <- yield17_w_pf %>%
  mapview(zcol = "yield_lbac",
        col.regions = viridis(3),
        alpha.regions = 1,
        cex = .1,
        lwd = 0,
        map.types = "Esri.WorldImagery"
)
yieldmap_clean_int
```

## Side-by-side raw vs. cleaned yield map  
```{r}
yieldmap_raw + yieldmap_clean
```

## Side-by-side raw vs. cleaned yield map interactive  
```{r}
sync(yieldmap_raw_int, yieldmap_clean_int)
```

# Exporting cleaned version  
```{r}
write_sf(yield17_w_pf,
         "../output/yield17_clean.geojson",
         delete_dsn = TRUE)

```

# Summary  
Today we:  

- Learned about and implemented 5 yield filters (yield, speed, position)  
- Compared raw and cleaned yield monitor data




